### 2016-12-19 doc/paper.tex
paper init

% Created 2016-12-16 Fr 23:12
% Intended LaTeX compiler: pdflatex
\documentclass[11pt]{llncs}

\tolerance=1000

\usepackage[T1]{fontenc}
%% % \usepackage[longnamesfirst,sort,nonamebreak]{natbib}
%% \usepackage[normalem]{ulem}
%% \usepackage[table]{xcolor}
\usepackage[utf8]{inputenc}
%% \usepackage{amsmath}
%% \usepackage{amssymb}
%% \usepackage{amstext}
%% \usepackage{array}
%% \usepackage{authblk}
%% \usepackage{caption}
%% \usepackage{color}
%% \usepackage{epsfig} % for postscript graphics files
%% \usepackage{fixltx2e}
%% \usepackage{float}
%% % \usepackage{fullpage}
%% \usepackage{graphics} % for pdf, bitmapped graphics files
\usepackage{hyperref}
%% \usepackage{latexsym}
%% \usepackage{lmodern}
%% \usepackage{longtable}
%% \usepackage{makeidx}  % allows for indexgeneration
%% \usepackage{marvosym}
%% \usepackage{multirow}
%% \usepackage{subcaption}
%% % \usepackage{subfigure}
%% \usepackage{textcomp}
%% \usepackage{titling}
%% \usepackage{wasysym}
%% \usepackage{wrapfig}

% \renewcommand{\familydefault}{\sfdefault}
% \parindent10pt

\usepackage[style=authoryear,backend=biber,bibencoding=utf8,natbib]{biblatex}
\bibliography{actinf}
\addbibresource{actinf.bib}

\usepackage{tikz}
\usetikzlibrary{arrows,positioning}

\graphicspath{{./img/}{img/}}
\DeclareGraphicsRule{.ps.gz}{eps}{.ps.bb}{`zcat #1}

\newcommand{\blg}[1]{
(\textcolor{red}{blg. #1})
}

%------------------------------------------------------------------------- 
% args: <label> <fileName> <caption> <scale>
\newcommand{\figone}[4]{
  \begin{figure}[ht]
  \begin{center}
  \includegraphics[scale=#4]{#2}
  \begin{sl}
  \caption{\label{#1}#3}
  \end{sl}
  \end{center}
  \end{figure}
}
%\figone{fig1}{figura.pdf}{caption de la figura}{5}

%------------------------------------------------------------------------- 
% args: <label> <fileName> <caption> <scale>
\newcommand{\figoneforced}[4]{
  \begin{figure}[h!t]
  \begin{center}
  \includegraphics[scale=#4]{#2}
  \begin{sl}
  \caption{\label{#1}#3}
  \end{sl}
  \end{center}
  \end{figure}
}
%\figone{fig1}{figura.pdf}{caption de la figura}{5}

%to insert a tikz fig in the local directory plus /figs
% args: <label> <fileName> <caption> 
\newcommand{\figtikz}[4]{
\begin{figure}[ht]
\begin{center}
\scalebox{#4}{\input{figs/#2}}
\begin{sl}
\caption{\label{#1}#3}
\end{sl}
\end{center}
\end{figure}
}
%-------------------------------------------------------------------------

%to insert a tikz fig in the local directory plus /figs
% args: <label> <fileName> <caption> acuerdate esta instruccion esta mal
\newcommand{\figtikzF}[4]{
\begin{figure}[h!t]
\begin{center}
\scalebox{#4}{\input{figs/#2}}
\begin{sl}
\caption{\label{#1}#3}
\end{sl}
\end{center}
\end{figure}
}
%-------------------------------------------------------------------------

% \title{Predictive Processing Model for an Autonomous Agent}
\title{Predictive Processing Model for an Autonomous Agent}
\titlerunning{Predictive Processing Model}  % abbreviated title (for running head)

\author{Bruno Lara\thanks{bruno.lara@uaem.mx}\inst{1} \and Alejandra
  Ciria \inst{2}
  \and Guido Schilacci \inst{3} \and Oswald Berthold \inst{3}}

%% \renewcommand\Authands{ and }

\institute{Cognitive Robotics Group, Center for Science Research,
  Universidad Aut\'onoma del Estado de Morelos, Cuernavaca, M\'exico
  \and Psychology Faculty, Universidad Nacional Aut\'onoma de
  M\'exico, M\'exico
  \and Adaptive Systems Group, Department of Computer Science,
  Humboldt-Universitaet zu Berlin, Berlin, Germany
}

% \author{Guido, Bruno, Alejandra, Esau, Oswald}
% \date{\today}
%% \hypersetup{
%%  pdfauthor={Guido, Bruno, Alejandra, Esau, Oswald},
%%  pdftitle={Predictive Processing Model for an Autonomous Agent},
%%  pdfkeywords={},
%%  pdfsubject={},
%%  pdfcreator={Emacs 24.3.1 (Org mode 9.0.1)}, 
%%  pdflang={English}}

\begin{document}

\maketitle

\begin{abstract}
We examine an approach to autonomous acquisition of motor skills and
behaviour using computational models. Our focus is on the role of
\emph{predictive coding} of sensorimotor data. Based on the apparent
prevalence of predictive coding in animal's brains, we follow the
hypothesis that predictive coding provides systematic and unique
characteristics facilitating learning and exploration.
% Short description of the project
\end{abstract}

\section{Introduction}
\label{intro}

Motor control for autonmous agents that need to learn some things
about the relations between observable quantities.

Our work is motivated by wanting to better understand how animals such as
ourselves are able to change, adjust and invent behaviour as
efficiently as they do. We assume such understanding to be relevant
for building a theory of robots, guidelines for the synthesis of
artificial systems interacting with the real world.

Existing research has shown the problem to be quite challenging and
varied and many such variations have been examined. The task is to try
and systematize the phenomena by identifying functional principles
that are at work over wide ranges of organisms and mechanisms
\citep{pfeifer07}.

We examine a small set of such principles in combination: Predictive
coding, internal modelling, and top-down information flow. We start
with the principle of Predicitive Coding (PC) which is a coding scheme
encountered in the analysis of nervous systems. The principle says
that of all input only that part of the input to a functional unit,
implemented by a layer or a module, which has not been predicted is
allowed to pass through for further processing. This quantity, usually
a vector, is the prediction error (PE).

The prediction error quantifies what the system doesn't know already
and is worthy of additional predictive modelling. PE combines external
(input complexity conditioned on output) and internal (model failures)
factors. Knowing the relation of model change to PE change allows to
infer particular sensorimotor contexts as driven by external
factors. Longterm PE behaviour can modulate adaptation, while
shortterm PE generally drives activity.

% ``inherent unpredictability''
% A possible hypothesis is .. damn it escaped me ..

Next is internal modelling (IM). IM is closely related to PE because
any prediction in fact requires a model. Such a model needs to map the
information from the past to the scope of prediction, usually one or a
few timesteps (1s).

Top-down information flow then says that experience, the apex of
decision making, is being driven or initiated by a prediction about
the next state. Only unpredicted (unmodelled) parts of the next
post-prediction measurements are being passed on.

%% Building on that and other existing concepts, the internal model
%% hypothesis (IMHT), and prediction / prediction error information flow
%% (PPEIF) we propose and analyze a simple model complying with the
%% principles.

We can take these as these three principles for the design of the
proposed models:

\begin{itemize}

\item Proprioceptive space: any embodied agent has to have a lowest
  level motor system. This is usually represented by a vector of motor
  angles, motor velocities, motor torques, but essentially be any
  given kind of controller. A controller in this sense implies a
  measurement and a prediction defined in the same space. This is the
  point at which we want to engange in interaction with the
  system.

\item The forward model is part of a larger building block, a
  generative model, mapping from control (goals) to to sensory
  consequences \citep{friston2011}.

\item The main flow of information is top-down, not bottom-up, and the
  forward flow of sensory information is replaced by the forward flow
  of prediction error \citep{clark2015}. FIXME: define top, botton,
  forward, etc?

\item Sensory prediction errors ($Err_{e(t)}$, $Err_{p(t)}$) are
  required for online state estimation (inference) and for optimizing
  (learning) the forward model \citep{friston2011}, or any kind of
  activity.

\item Descending propioceptive predictions are equivalent to motor
  commands. Movement quash error at the level of spinal reflexes
  \citep{pickering2014}. Propioceptive predictions elicit motor
  actions, so motor commands, so motor commands are replaced by those
  propioceptive predictions \citep{clark2015}. $S^*_{p(t)}=M_{(t)}$.

\item Motor commands become descending predictions or propioceptive
  sensations, while their exteroceptive homologous become corollary
  discharges \citep{friston2011}.

\item Motor behaviours are predictions of propioceptive
  patterns. Motor control is just more top-down sensory prediction
  \citep{clark2015}.

\item Action: results from our own predictions concerning the flow of
  sensations \citep{pickering2014}.

\item High-level plans and intentions (prior beliefs) associate with
  sensory consequences and actions flow. This is, high level states
  into predicted sensory patterns \citep{pickering2014}.

\end{itemize}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Related}
\label{sec:orgf9aed21}

Long history, many models: control, development, RL, \ldots{}

\section{Proposed Models}
\label{plan}

\subsection{General Models}

This model includes both propio and extero (Figs \ref{fig:basic1}
and \ref{fig:basic2}). The only difference is how the 
predictive error is calculated for the propioceptive information ($Err_{Pi(t)}$).

%\figtikz{fig:ltp}{fig_im_ltp}{Long term prediction}{0.9}

\figone{fig:basic1}{Integral_v3}{Model A}{0.8}
\figone{fig:basic2}{Integral_v4}{Model B}{0.8}

\subsection{Case Study Models}

In the example to simulate, we use the visual modality as exteroceptive
information.

The flow of the experiment, once trained would be as follows:

\begin{enumerate}
\item The visual starting position of the hand is located in space and becomes $V_i$
\item A visual target is located and becomes $V_{goal}$.
\item This two visual positions in space are mapped to the  $MMR$ map which
  codes for trajectories in the 2D space of the robot. For each pair of
  positions in visual space exists a pair of nodes in motor space. This can be
  seen as previous knowledge  and does the required mapping between the two
  spaces. A version of this is already implemented. (see Section \ref{subsec:exteropropio})
\item With these positions in both spaces, the flow of the experiment goes
  down to the forward models. 
\item Movement in the intrinsic frame of reference, as already implemented in the propioceptive model, has
  consequences on the environment which cause changes in the extrinsic frame
  of reference.

\end{enumerate}

We can talk of two different versions of this, as we can use directly just the
initial and goal situations in both maps or in a more precise way use
intermediate points to perform micro-predictions to correct the trajectory.

The two models shown have the same difference as the previous one, this is,
the use of the prediction for obtaining the predictive error.

\begin{figure}
  \includegraphics[width=0.99\textwidth]{img/actinf_large_diagram_p_e2p_e.pdf}
  \caption{Hello grpahicxs}
\end{figure}

\subsection{Propio-Extero Mapping Model}
\label{subsec:exteropropio}

There would be here, the part that Guido was implementing on the forming of trajectories.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Notes}
\label{sec:org863d8de}

\subsection{control theory parallels}
\label{sec:org316353f}

i don't see this as a problem at all, to the contrary, it means if you
arrive at a model setup you can actually look to adaptive control and
translate the solution to your formulation.

in terms of the philosophy see email from \textit{<2016-10-11 Di> } so it's
clearly about a different issue.

\subsection{merge bruno's initial draft}
\label{sec:org6e00d89}

\subsection{opt's notes}
\label{sec:org35938d1}
\begin{itemize}
\item for proprio-only show 3 figures: 1) behaviour without any learning,
which fails to precisely reach the goals because proprio
predictions don't "match" proprio states (nonlinear
deformations), 2) behaviour during learning, see how we actually
move closer to the goal over a few timesteps, 3) behaviour while
still learning but going to places we have already been: no error,
no learning / weight changes
\item additional plots: original timeseries, model sweep response plot
\end{itemize}

%% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% \section{Introduction}
%% \label{sec:orga5a5ee2}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Core idea and models}
\label{sec:org733d6e8}
\begin{enumerate}
\item discuss the change of terminology or dichotomies regarding:
\begin{itemize}
\item predictions/motor commands,
\item forward models/inverse models/controllers/policies
\item bottom level black box: different actual forward model (FM)
representations via SOMs, kNN, RLS, GP, \ldots{}
\item see discussion with antonio re: "it's just labels", underneath it is a raw
network of signal flows putting observable quantities into
relation to each other, from which apparent goal-driven behaviour
(teleology) emerges
\end{itemize}
\item two models and their difference in terms of learning dynamics. do
any dynamic properties support or contradict biological motor learning?
\begin{enumerate}
\item M1: explicit goal + error as input
\item M2: error only as input
\item merge with alejandra's two versions (explicit update rule)
\end{enumerate}
\item Experiments
\begin{itemize}
\item deploy models to different robotic systems: Nao, SimpleArm,
Pointmass, two-wheeled
\item timeseries analysis
\item learning transient (arbitrarily fast, see Tin + Poon)
\item explicitly map out the acquired models by sweeping the inputs and
record their outputs
\item what happens under perturbation?
\item long-term adaptation effects: are we only plastic or are we also
stable? do new adaptations overwrite previous ones? does a given
approximator become saturated (just think how critical weight
initialization in neural networks can be) in terms of its adaption
capacity?, can be enforced via decay schedules of learning rate
etc
\end{itemize}
\item discuss, compare with existing approaches like std fwd/inv model
pairs or reinforcement learning
\end{enumerate}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Experiments}
\label{sec:org347db7c}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Discussion}
\label{sec:orga3d6260}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Summary}
\label{sec:orgc71dbe7}

\printbibliography
\end{document}

### 2016-12-20 doc/paper.tex
papr onwards

@@
 
-We can take these as these three principles for the design of the
-proposed models:
+We can take these three principles for the design of the proposed models:
 
@@
 
+\item Discuss the change of terminology or dichotomies regarding:
+  \begin{itemize}
+  \item predictions/motor commands,
+  \item forward models/inverse models/controllers/policies
+  \item bottom level black box: different actual forward model (FM)
+    representations via SOMs, kNN, RLS, GP, \ldots{}
+  \item see discussion with antonio re: "it's just labels", underneath it is a raw
+    network of signal flows putting observable quantities into
+    relation to each other, from which apparent goal-driven behaviour
+    (teleology) emerges
+  \end{itemize}
+
 \item Proprioceptive space: any embodied agent has to have a lowest
   level motor system. This is usually represented by a vector of motor
-  angles, motor velocities, motor torques, but essentially be any
+  angles, motor velocities, motor torques, etc but essentially it can be any
   given kind of controller. A controller in this sense implies a
@@
   \citep{pickering2014}. Propioceptive predictions elicit motor
-  actions, so motor commands, so motor commands are replaced by those
+  actions, so motor commands are replaced by those
   propioceptive predictions \citep{clark2015}. $S^*_{p(t)}=M_{(t)}$.
@@
 
-Long history, many models: control, development, RL, \ldots{}
+Long history, many models: control, development, RL, neuroscience, ...\ldots{}
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Proposed Models}
@@
 \subsection{General Models}
+
+First two basic models and their difference in terms of learning
+dynamics. Do any dynamic properties support or contradict biological
+motor learning?
+
+\subsection{M1}
+
+The basic building block (BBB) contains an adaptive forward model
+(predictor) whose output is the BBB's output. We assume direct
+feedback from that module that the output is connected to. The BBB'
+inputs are an explicit goal and the prediction error with respect to
+the last prediction at ${t-1}$. These inputs taken together encode the
+state. Thus the forward model should in principle be able to pick up
+on different relations depending on the state. Graphical
+representation is given in \autoref{fig:M1}.
+
+\begin{figure}
+  \centering
+  \includegraphics[width=0.9\textwidth]{img/mdl_prediction_error_and_goal.pdf}
+  \caption{\label{fig:M1} Basic model M1.}
+\end{figure}
+
+\subsection{M2}
+
+This modification of BBB M1 receives only the error as an input. The
+model's graph is shown in \autoref{fig:M2}. This model will not be
+able to distinguish between situations demanding different
+predictions.
+
+\begin{figure}
+  \centering
+  \includegraphics[width=0.9\textwidth]{img/mdl_prediction_error_only.pdf}
+  \caption{\label{fig:M2} Basic model M2.}
+\end{figure}
+
+\subsection{Extended models connecting modalities}
+
+Merge with alejandra's two versions (explicit update rule).
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Core idea and models}
-\label{sec:org733d6e8}
-\begin{enumerate}
-\item discuss the change of terminology or dichotomies regarding:
-\begin{itemize}
-\item predictions/motor commands,
-\item forward models/inverse models/controllers/policies
-\item bottom level black box: different actual forward model (FM)
-representations via SOMs, kNN, RLS, GP, \ldots{}
-\item see discussion with antonio re: "it's just labels", underneath it is a raw
-network of signal flows putting observable quantities into
-relation to each other, from which apparent goal-driven behaviour
-(teleology) emerges
-\end{itemize}
-\item two models and their difference in terms of learning dynamics. do
-any dynamic properties support or contradict biological motor learning?
-\begin{enumerate}
-\item M1: explicit goal + error as input
-\item M2: error only as input
-\item merge with alejandra's two versions (explicit update rule)
-\end{enumerate}
-\item Experiments
+\section{Experiments}
+\label{sec:org347db7c}
+
   \begin{itemize}
@@
   \end{itemize}
-\item discuss, compare with existing approaches like std fwd/inv model
-pairs or reinforcement learning
-\end{enumerate}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Experiments}
-\label{sec:org347db7c}
 
@@
 \label{sec:orga3d6260}
+
+Discuss, compare with existing approaches like std fwd/inv model
+  pairs or reinforcement learning
 

### 2016-12-20 actinf_models.py
using pointmass1d/3d as alternative system

@@ -103,8 +103,8 @@ class ActInfKNN(ActInfModel):
         print("%s.bootstrap'ping" % (self.__class__.__name__))
         for i in range(10):
             if self.idim == self.odim:
-                self.X_.append(np.ones((self.idim, ) * i * 0.1))
-                self.y_.append(np.ones((self.odim, ) * i * 0.1))
+                self.X_.append(np.ones((self.idim, )) * i * 0.1)
+                self.y_.append(np.ones((self.odim, )) * i * 0.1)
             else:
                 self.X_.append(np.random.uniform(-0.1, 0.1, (self.idim,)))
                 self.y_.append(np.random.uniform(-0.1, 0.1, (self.odim,)))

### 2016-12-20 active_inference_basic.py

@@ -53,6 +53,7 @@ modes = [
 class ActiveInferenceExperiment(object):
     def __init__(self, mode = "type01_state_prediction_error",
                  model = "knn", numsteps = 1000, idim = None,
+                 environment_str = "simplearm",
                  goal_sample_interval = 50, e2pmodel = None):
         self.mode = mode
         self.model = model
@@ -62,18 +63,47 @@ class ActiveInferenceExperiment(object):
         # experiment settings
         self.numsteps = numsteps
         
-        # intialize robot environment        
-        self.environment = Environment.from_configuration('simple_arm', 'low_dimensional')
-        # self.environment = Environment.from_configuration('pointmass', 'low_dim_vel')
-        self.environment.noise = 0.
-
-        if idim is None:
-            self.idim = self.environment.conf.m_ndims * 2
+        # intialize robot environment
+        if environment_str == "simplearm":
+            self.environment = Environment.from_configuration('simple_arm', 'low_dimensional')
+            self.environment.noise = 0.
+            # dimensions
+            if mode.startswith("type03_1"):
+                self.idim = self.environment.conf.m_ndims
+            else:
+                self.idim = self.environment.conf.m_ndims * 2
+            self.odim = self.environment.conf.m_ndims
+            self.dim_ext  = 2 # cartesian
+                
+        elif environment_str == "pointmass1d":
+            self.environment = Environment.from_configuration('pointmass', 'low_dim_vel')
+            if mode.startswith("type03_1"):
+                self.idim = self.environment.conf.m_ndims
+            else:
+                self.idim = self.environment.conf.m_ndims * 2
+            self.odim = self.environment.conf.m_ndims
+            self.dim_ext  = self.environment.conf.m_ndims # cartesian
+            
+        elif environment_str == "pointmass3d":
+            self.environment = Environment.from_configuration('pointmass', 'mid_dim_vel')
+            if mode.startswith("type03_1"):
+                self.idim = self.environment.conf.m_ndims
+            else:
+                self.idim = self.environment.conf.m_ndims * 2
+            self.odim = self.environment.conf.m_ndims
+            self.dim_ext  = self.environment.conf.m_ndims # cartesian
         else:
-            self.idim = idim
[…]

### 2016-12-20 doc/paper.tex
using pointmass1d/3d as alternative system
preparing plots for section (currently) 5.1 example M1

@@
 %% \usepackage{latexsym}
-%% \usepackage{lmodern}
+\usepackage{lmodern}
 %% \usepackage{longtable}
@@
 %% \usepackage{subcaption}
-%% % \usepackage{subfigure}
+\usepackage[tight,footnotesize]{subfigure}
 %% \usepackage{textcomp}
@@
 \bibliography{actinf}
-\addbibresource{actinf.bib}
+% \addbibresource{actinf.bib}
 
@@
 
-Long history, many models: control, development, RL, neuroscience, ...\ldots{}
+Long history, many models: control, development, RL, neuroscience, prediction and prediction error (residual), ...\ldots{}
 
@@
 
-\subsection{General Models}
+\subsection{Basis building blocks}
 
@@
 (predictor) whose output is the BBB's output. We assume direct
-feedback from that module that the output is connected to. The BBB'
+feedback from the module that the output is connected to. The BBB'
 inputs are an explicit goal and the prediction error with respect to
@@
 on different relations depending on the state. Graphical
-representation is given in \autoref{fig:M1}.
+representation is given in \autoref{fig:M1}. The closed-loop dynamics
+of this model are described in detail in \autoref{sec:exprs_basic_ex_m1}.
+
+FIXME: Fundamental limitation: can only learn monotonic relationships so
+far. Show to fail, indicate amendment that can solve non-monotonic
+cases (by observing PE over time, by using reward modulation, ...).
 
@@
 able to distinguish between situations demanding different
-predictions.
+predictions. FIXME: It can be shown to fail when such a situation is produced
+by experiment design.
 
@@
 \end{figure}
+
+\subsection{E2P}
+
+A third building block is the intermodal map, specifically mapping
+between an exteroceptive and the proprioceptive modality (thus
+E2P). We propose to realize this slightly differently from M1 /
+M2. The E2P module is bootstrapped by passively observing
+corresponding exteroceptive and proprioceptive data and learning those
+correspondences. Once bootstrapped, the E2P can be used to predict
+possibly ambigous proprioceptive configurations from the exeroceptive
+sensations they are believed to produce, meaning which have been
+observed in the past to do so.
 
@@
 
+Notes for experiments:
+
 \begin{itemize}
@@
 
+\subsection{Basic behavioural example M1}
+\label{sec:exprs_basic_ex_m1}
+
+\begin{itemize}
+\item Basic operation:
+  \autoref{fig:basic_op_200_knn}, \autoref{fig:basic_op_200_soesgp},
+  \autoref{fig:basic_op_1000_knn}, \autoref{fig:basic_op_1000_soesgp}
+  show variations on the basic operation theme. We are evaluating a
+  k-NN and a Gaussian Process based forward model each over 200 and
+  1000 timesteps interaction on a one-dimensional identity system. The
+  forward model starts off in a zero-state, the first prediction being
+  zero. Any resulting prediction error drives an adaptation on the
+  forward model output towards reduced error for a given goal, error
+  input pair. We sweep this
+  input space once at numsteps/2 and once at the end of the
+  experiment which demonstrates the change in the overall model
+  response.
+  \item Compare identity system with small aberration system: now we
+    can do the same for a 1D system while introducing aberrations in
+    the system's response function.
+  \item Failure under non-monotonicity: if the aberrations become
+    systemically non-monotonic the learning process fails (FIXME: at least I
+    predict that ;)
+\item How PE is low when revisiting previously observed
+  configurations: longterm PE plot of 10000 timesteps and averaged PE
+  in addition to instantaneous PE
+% \item operation using non-identity motor transfer function
+\end{itemize}
+
+\begin{figure}
+  \begin{center}
+  \subfigure[System sweep response]{
+    \includegraphics[width=0.47\textwidth]{img/basic_op_id_200steps_knn/fig_001_scattermatrix_reduced.pdf}%
+    \label{fig:basic_op_200_knn_fig1}
+  }
+  \subfigure[Goal, prediction, error]{
+    \includegraphics[width=0.47\textwidth]{img/basic_op_id_200steps_knn/fig_003_aie_experiment_plot_basic.pdf}
+    \label{fig:basic_op_200_knn_fig2}
+  }
+  \subfigure[Model sweep t = 100]{
+    \includegraphics[width=0.47\textwidth]{img/basic_op_id_200steps_knn/fig_002_colormeshmatrix_reduced.pdf}
+    \label{fig:basic_op_200_knn_fig3}
+  }
+  \subfigure[Model sweep t = 200]{
+    \includegraphics[width=0.47\textwidth]{img/basic_op_id_200steps_knn/fig_004_colormeshmatrix_reduced.pdf}
+    \label{fig:basic_op_200_knn_fig4}
+  }
+  \end{center}
+  \caption{\label{fig:basic_op_200_knn}Basic operation 200 timesteps knn.}
+\end{figure}
+
+\begin{figure}
+  \begin{center}
+  \subfigure[System sweep response]{
+    \includegraphics[width=0.47\textwidth]{img/basic_op_id_200steps_soesgp/fig_001_scattermatrix_reduced.pdf}%
+    \label{fig:basic_op_200_soesgp_fig1}
+  }
+  \subfigure[Goal, prediction, error]{
+    \includegraphics[width=0.47\textwidth]{img/basic_op_id_200steps_soesgp/fig_003_aie_experiment_plot_basic.pdf}
+    \label{fig:basic_op_200_soesgp_fig2}
+  }
+  \subfigure[Model sweep t = 100]{
+    \includegraphics[width=0.47\textwidth]{img/basic_op_id_200steps_soesgp/fig_002_colormeshmatrix_reduced.pdf}
+    \label{fig:basic_op_200_soesgp_fig3}
+  }
+  \subfigure[Model sweep t = 200]{
+    \includegraphics[width=0.47\textwidth]{img/basic_op_id_200steps_soesgp/fig_004_colormeshmatrix_reduced.pdf}
+    \label{fig:basic_op_200_soesgp_fig4}
+  }
+  \end{center}
+  \caption{\label{fig:basic_op_200_soesgp}Basic operation 200 timesteps GP.}
+
+\end{figure}
+\begin{figure}
+  \begin{center}
+  \subfigure[System sweep response]{
+    \includegraphics[width=0.47\textwidth]{img/basic_op_id_1000steps_knn/fig_001_scattermatrix_reduced.pdf}%
+    \label{fig:basic_op_1000_knn_fig1}
+  }
+  \subfigure[Goal, prediction, error]{
+    \includegraphics[width=0.47\textwidth]{img/basic_op_id_1000steps_knn/fig_003_aie_experiment_plot_basic.pdf}
+    \label{fig:basic_op_1000_knn_fig2}
+  }
+  \subfigure[Model sweep t = 500]{
+    \includegraphics[width=0.47\textwidth]{img/basic_op_id_1000steps_knn/fig_002_colormeshmatrix_reduced.pdf}
+    \label{fig:basic_op_1000_knn_fig3}
+  }
+  \subfigure[Model sweep t = 1000]{
+    \includegraphics[width=0.47\textwidth]{img/basic_op_id_1000steps_knn/fig_004_colormeshmatrix_reduced.pdf}
+    \label{fig:basic_op_1000_knn_fig4}
+  }
+  \end{center}
+  \caption{\label{fig:basic_op_1000_knn}Basic operation 1000 timesteps knn.}
+\end{figure}
+
+\begin{figure}
+  \begin{center}
+  \subfigure[System sweep response]{
+    \includegraphics[width=0.47\textwidth]{img/basic_op_id_1000steps_soesgp/fig_001_scattermatrix_reduced.pdf}%
+    \label{fig:basic_op_1000_soesgp_fig1}
+  }
+  \subfigure[Goal, prediction, error]{
+    \includegraphics[width=0.47\textwidth]{img/basic_op_id_1000steps_soesgp/fig_003_aie_experiment_plot_basic.pdf}
+    \label{fig:basic_op_1000_soesgp_fig2}
+  }
+  \subfigure[Model sweep t = 500]{
+    \includegraphics[width=0.47\textwidth]{img/basic_op_id_1000steps_soesgp/fig_002_colormeshmatrix_reduced.pdf}
+    \label{fig:basic_op_1000_soesgp_fig3}
+  }
+  \subfigure[Model sweep t = 1000]{
+    \includegraphics[width=0.47\textwidth]{img/basic_op_id_1000steps_soesgp/fig_004_colormeshmatrix_reduced.pdf}
+    \label{fig:basic_op_1000_soesgp_fig4}
+  }
+  \end{center}
+  \caption{\label{fig:basic_op_1000_soesgp}Basic operation 1000 timesteps GP.}
+
+\end{figure}
+
+\subsection{Basic behavioural example M2}
+\label{sec:exprs_basic_ex_m2}
+\subsection{Basic behavioural example E2P}
+\label{sec:exprs_basic_ex_e2p}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2016-12-20 active_inference_basic.py

@@ -1,10 +1,12 @@
 
 import argparse, cPickle, os
 from collections import OrderedDict
-from functools   import partial
+from functools   import partial # mhm :)
 
 import numpy as np
 import pylab as pl
+import matplotlib.gridspec as gridspec
+import pandas as pd
 
 from explauto import Environment
 from explauto.environment import environments
@@ -28,7 +30,12 @@ try:
 except ImportError, e:
     print "Couldn't import pypr.clustering.gmm", e
 
-# TODO: modify this code to use the GMM model from actinf_models.py
+SAVEPLOTS = True
+    
+# TODO: make fwd model aware of goal change to ignore old prediction error which is irrelevant for the new goal
+# TODO: make fwd model aware of it's learning state and wether it can advantageously accomodate new data or
+#       is fine to ignore it. Also: prune old data?
+    
 # TODO: variant 0 and 1: multiply error with (the sign of) local gradient of the function you're modelling for non-monotonic cases?
 #       need to plot behaviour for non-invertible functions
 # TODO: make custom models: do incremental fit and store history for
@@ -39,6 +46,8 @@ except ImportError, e:
 # TODO: compute PI,AIS,TE for goal->state, input->pred_error, s_pred->s to
 #       answer qu's like: how much is there to learn, how much is learned
 # TODO: pass pre-generated system into actinf machine, so we can use random robots with parameterized DoF, stiffness, force budget, DoF coupling coefficient
+
+# DONE: modify this code to use the GMM model from actinf_models.py (ca. 20161210)
 
 modes = [
     "test_models",
@@ -54,7 +63,8 @@ class ActiveInferenceExperiment(object):
     def __init__(self, mode = "type01_state_prediction_error",
                  model = "knn", numsteps = 1000, idim = None,
                  environment_str = "simplearm",
-                 goal_sample_interval = 50, e2pmodel = None):
+                 goal_sample_interval = 50, e2pmodel = None,
+                 saveplots = False):
         self.mode = mode
         self.model = model
 
@@ -62,11 +72,13 @@ class ActiveInferenceExperiment(object):
 
         # experiment settings
         self.numsteps = numsteps
+
+        self.saveplots = saveplots
         
         # intialize robot environment
         if environment_str == "simplearm":
             self.environment = Environment.from_configuration('simple_arm', 'low_dimensional')
-            self.environment.noise = 0.
[…]

### 2016-12-21 actinf_models.py
redone 1d model plots for 5.1 M1 to equalize color normalization

@@ -1316,6 +1316,9 @@ def generate_inverted_sinewave_dataset(N = 1000):
 
 def test_model(args):
     import pylab as pl
+
+    np.random.seed(args.seed)
+    
     # get last component of datafile, the actual filename
     datafilepath_comps = args.datafile.split("/")
     if datafilepath_comps[-1].startswith("EP"):
@@ -1432,10 +1435,11 @@ if __name__ == "__main__":
     import argparse
 
     parser = argparse.ArgumentParser()
-    parser.add_argument("-d", "--datafile",  type=str, help="datafile containing t x (dim_extero + dim_proprio) matrix ", default="data/simplearm_n1000/EP_1000.npy")
-    parser.add_argument("-m", "--modelclass", type=str, help="Which model class to test [KNN], " + ", ".join(model_classes), default="KNN")
-    parser.add_argument("-n", "--numsteps",  type=int, help="Number of datapoints [1000]", default=1000)
-    parser.add_argument("-ne", "--numepisodes",  type=int, help="Number of episodes [10]", default=10)
+    parser.add_argument("-d", "--datafile",     type=str, help="datafile containing t x (dim_extero + dim_proprio) matrix ", default="data/simplearm_n1000/EP_1000.npy")
+    parser.add_argument("-m", "--modelclass",   type=str, help="Which model class to test [KNN], " + ", ".join(model_classes), default="KNN")
+    parser.add_argument("-n", "--numsteps",     type=int, help="Number of datapoints [1000]", default=1000)
+    parser.add_argument("-ne", "--numepisodes", type=int, help="Number of episodes [10]", default=10)
+    parser.add_argument("-s",  "--seed",        type=int, help="seed for RNG [0]",        default=0)
     args = parser.parse_args()
     
     test_model(args)

### 2016-12-21 active_inference_basic.py

@@ -916,7 +916,7 @@ class ActiveInferenceExperiment(object):
  
         # plot_scattermatrix(df)
         # plot_scattermatrix_reduced(df)
-        plot_colormeshmatrix_reduced(self.X_model_sweep, self.Y_model_sweep)
+        plot_colormeshmatrix_reduced(self.X_model_sweep, self.Y_model_sweep, ymin = -1.0, ymax = 1.0)
         
     ################################################################################
     def map_model_type03_goal_prediction_error(self):
@@ -1430,7 +1430,7 @@ def plot_scattermatrix_reduced(df):
         fig.savefig("fig_%03d_scattermatrix_reduced.pdf" % (fig.number), dpi=300)
     fig.show()
             
-def plot_colormeshmatrix_reduced(X, Y):
+def plot_colormeshmatrix_reduced(X, Y, ymin = None, ymax = None):
     print "X.shape", X.shape, "Y.shape", Y.shape
     # input_cols  = [i for i in df.columns if i.startswith("X")]
     # output_cols = [i for i in df.columns if i.startswith("Y")]
@@ -1442,7 +1442,6 @@ def plot_colormeshmatrix_reduced(X, Y):
     
     # # numplots = Xs.shape[1] * Ys.shape[1]
     # # print "numplots = %d" % numplots
-
     
     gs = gridspec.GridSpec(Y.shape[2], X.shape[2]/2)
     pl.ioff()
@@ -1455,7 +1454,7 @@ def plot_colormeshmatrix_reduced(X, Y):
         for j in range(Y.shape[2]):
             # print "i, j", i, j, Xs, Ys
             ax = fig.add_subplot(gs[j, i])
-            pcm = ax.pcolormesh(X[:,:,0], X[:,:,1], Y[:,:,0])
+            pcm = ax.pcolormesh(X[:,:,0], X[:,:,1], Y[:,:,0], vmin = ymin, vmax = ymax)
             # ax.plot(Xs.as_matrix()[:,i], Ys.as_matrix()[:,j], "ko", alpha = alpha)
             ax.set_xlabel("goal")
             ax.set_ylabel("error")
@@ -1488,6 +1487,10 @@ def test_models(args):
 
     
 def main(args):
+
+    # seed PRNG
+    np.random.seed(args.seed)
+        
     if args.mode.startswith("test_"):
         test_models(args)
     else:
@@ -1515,6 +1518,7 @@ if __name__ == "__main__":
[…]

### 2016-12-21 active_inference_basic.py
going wild on aberrations and non-monotonicity

@@ -73,8 +73,12 @@ class ActiveInferenceExperiment(object):
         # experiment settings
         self.numsteps = numsteps
 
+        self.eta_fwd_mdl = 0.7
+        self.coef_smooth_fast = 0.9
+        self.coef_smooth_slow = 0.95
+                
         self.saveplots = saveplots
-        
+
         # intialize robot environment
         if environment_str == "simplearm":
             self.environment = Environment.from_configuration('simple_arm', 'low_dimensional')
@@ -137,18 +141,23 @@ class ActiveInferenceExperiment(object):
         # logging, configure logs dictionary and dict of logging variables
         self.logs = {}
         self.logging_vars = {
-            "S_prop_pred": {"col": self.odim},
-            "E_prop_pred": {"col": self.odim},
-            "M_prop_pred": {"col": self.odim},
-            "goal_prop":   {"col": self.odim},
-            "S_ext":       {"col": self.dim_ext},
-            "E2P_pred":    {"col": self.odim},
-            "P2E_pred":    {"col": self.dim_ext},
-            "goal_ext":    {"col": self.dim_ext},
-            "E_pred_e":    {"col": self.dim_ext},
-            "X_":          {"col": -1}, # 6
-            "X__":         {"col": -1}, # 6
-            "y_":          {"col": -1}  # 3 low-dim
+            "S_prop_pred":  {"col": self.odim},
+            "E_prop_pred":  {"col": self.odim},
+            "E_prop_pred_fast": {"col": self.odim},
+            "dE_prop_pred_fast": {"col": self.odim},
+            "E_prop_pred_slow": {"col": self.odim},
+            "dE_prop_pred_slow": {"col": self.odim},
+            "d_E_prop_pred_": {"col": self.odim},
+            "M_prop_pred":  {"col": self.odim},
+            "goal_prop":    {"col": self.odim},
+            "S_ext":        {"col": self.dim_ext},
+            "E2P_pred":     {"col": self.odim},
+            "P2E_pred":     {"col": self.dim_ext},
+            "goal_ext":     {"col": self.dim_ext},
+            "E_pred_e":     {"col": self.dim_ext},
+            "X_":           {"col": -1}, # 6
+            "X__":          {"col": -1}, # 6
+            "y_":           {"col": -1}  # 3 low-dim
[…]

### 2016-12-21 active_inference_basic.py
new model m2 using gradient estimation

@@ -50,11 +50,12 @@ SAVEPLOTS = True
 # DONE: modify this code to use the GMM model from actinf_models.py (ca. 20161210)
 
 modes = [
+    "m2_prediction_errors",
     "test_models",
     "type01_state_prediction_error",
     "type02_state",
-    "type03_goal_prediction_error",
     "type03_1_prediction_error",
+    "type03_goal_prediction_error",
     "type04_ext_prop",
     "type05_multiple_models",
 ]
@@ -260,6 +261,30 @@ class ActiveInferenceExperiment(object):
         elif mode == "type05_multiple_models":
             self.run = self.run_type05_multiple_models
 
+        elif mode == "m2_prediction_errors":
+            """active inference / predictive coding: first working, most basic version,
+            proprioceptive only
+
+            goal -> goal state prediction -> goal/state error -> update forward model"""
+
+            # pimp environment
+            self.environment.motor_aberration["type"] = "sin"
+            self.environment.motor_aberration["coef"] = -0.7
+            
+            # learning loop hooks
+            self.rh_learn_proprio_hooks["hook01"] = self.lh_learn_proprio_base_2
+            self.rh_learn_proprio_hooks["hook02"] = self.lh_learn_proprio_e2p2e
+            self.rh_learn_proprio_hooks["hook03"] = self.lh_do_logging
+            self.rh_learn_proprio_hooks["hook04"] = self.lh_sample_discrete_uniform_goal
+            # self.rh_learn_proprio_hooks["hook05"] = self.lh_sample_error_gradient
+
+            # experiment loop hooks
+            self.run_hooks["hook00"] = self.make_plot_system_function_and_exec # sweep system before learning
+            self.run_hooks["hook01"] = partial(self.rh_learn_proprio, iter_start = 0, iter_end = self.numsteps)
+            self.run_hooks["hook02"] = self.rh_learn_proprio_save
+            # self.run_hooks["hook03"] = self.rh_check_for_model_and_map
+            self.run_hooks["hook99"] = self.experiment_plot
+            
         elif mode == "basic_operation_1D_M1":
             """Experiment: Basic operation M1 on 1-dimensional data"""
 
@@ -459,12 +484,12 @@ class ActiveInferenceExperiment(object):
                 v(i)
 
[…]

### 2016-12-22 active_inference_basic.py
added pm3d model sweep, wip

@@ -269,13 +269,13 @@ class ActiveInferenceExperiment(object):
 
             # pimp environment
             self.environment.motor_aberration["type"] = "sin"
-            self.environment.motor_aberration["coef"] = -0.7
+            self.environment.motor_aberration["coef"] = np.random.uniform(-1.0, 1.0, self.odim) # -0.7
             
             # learning loop hooks
+            self.rh_learn_proprio_hooks["hook04"] = self.lh_sample_discrete_uniform_goal
             self.rh_learn_proprio_hooks["hook01"] = self.lh_learn_proprio_base_2
             self.rh_learn_proprio_hooks["hook02"] = self.lh_learn_proprio_e2p2e
             self.rh_learn_proprio_hooks["hook03"] = self.lh_do_logging
-            self.rh_learn_proprio_hooks["hook04"] = self.lh_sample_discrete_uniform_goal
             # self.rh_learn_proprio_hooks["hook05"] = self.lh_sample_error_gradient
 
             # experiment loop hooks
@@ -283,7 +283,8 @@ class ActiveInferenceExperiment(object):
             self.run_hooks["hook01"] = partial(self.rh_learn_proprio, iter_start = 0, iter_end = self.numsteps)
             self.run_hooks["hook02"] = self.rh_learn_proprio_save
             # self.run_hooks["hook03"] = self.rh_check_for_model_and_map
-            self.run_hooks["hook99"] = self.experiment_plot
+            self.run_hooks["hook98"] = self.experiment_plot
+            self.run_hooks["hook99"] = self.make_plot_model_function_and_exec # sweep model after learning
             
         elif mode == "basic_operation_1D_M1":
             """Experiment: Basic operation M1 on 1-dimensional data"""
@@ -418,7 +419,7 @@ class ActiveInferenceExperiment(object):
     def rh_system_generate_sweep_input(self):
         """generate system inputs for grid sweep"""
         # create meshgrid over proprio dimensions
-        sweepsteps = 21
+        sweepsteps = 11 # 21
         dim_axes = [np.linspace(self.environment.conf.m_mins[i], self.environment.conf.m_maxs[i], sweepsteps) for i in range(self.environment.conf.m_ndims)]
         full_axes = np.meshgrid(*tuple(dim_axes), indexing='ij')
 
@@ -571,7 +572,8 @@ class ActiveInferenceExperiment(object):
[…]

### 2017-06-06 actinf_models.py
trying to repair map_model before merging with smp_graphs

@@ -444,6 +444,7 @@ class ActInfGMM(ActInfModel):
                 self.cond = X[(i+ref_interval) % numsamplesteps] # self.X__[i,:3]
                 # self.cond = np.array()
                 # self.cond[:2] = X_
+                # print(self.cond, out_dims, X.shape)
                 self.cond[out_dims] = np.nan
                 (self.cen_con, self.cov_con, self.new_p_k) = gmm.cond_dist(self.cond, self.cen_lst, self.cov_lst, self.p_k)
                 # print "run_hook_e2p_sample gmm.cond_dist:", np.array(self.cen_con).shape, np.array(self.cov_con).shape, self.new_p_k.shape

### 2017-06-06 active_inference_basic.py

@@ -60,6 +60,12 @@ modes = [
     "type05_multiple_models",
 ]
 
+actinf_environments = [
+    'simplearm',
+    'pointmass1d',
+    'pointmass3d',
+]
+    
 def gaussian(m, s, x):
     """univariate gaussian"""
     return 1/(s*np.sqrt(2*np.pi)) * np.exp(-0.5*np.square((m-x)/s))
@@ -117,6 +123,8 @@ class ActiveInferenceExperiment(object):
             print "%s.__init__ unknown environment string '%s', exiting" % (self.__class__.__name__, environment_str)
             import sys
             sys.exit(1)
+
+        print "%s init pass 1: enironment = %s / %s, idim = %d, odim = %d" % (self.__class__.__name__, environment_str, self.environment, self.idim, self.odim)
 
         # if idim is None:
         #     self.idim = self.environment.conf.m_ndims * 2
@@ -189,27 +197,51 @@ class ActiveInferenceExperiment(object):
         self.S_prop_pred = np.random.normal(0, 0.01, (1, self.odim))
 
     def init_e2p(self, e2pmodel):
+        """ActiveInferenceExperiment.init_e2p
+
+        Initialize the extero-to-proprio mapping
+        """
+        # sum up all variable dimensions
         mmdims = self.dim_ext + self.odim
-        if e2pmodel == "gmm":
+        if e2pmodel == "gmm": # use gaussian mixture model
             self.mm = ActInfGMM(idim = self.dim_ext, odim = self.odim)
-        elif e2pmodel == "som":
+        elif e2pmodel == "som": # use hebbian SOM model
             self.mm = ActInfHebbianSOM(idim = self.dim_ext, odim = self.odim)
         else:
             print "unknown e2pmodel %s" % e2pmodel
         
     def init_wiring(self, mode):
-        """initialize structure of experiment set by 'mode'"""
+        """ActiveInferenceExperiment.init_wiring
+
+        Initialize the structure of the experiment specified by the 'mode'.
+
+        The experiment consists of a 'run' on the outside that executes
+        all functions registered in self.run_hooks once and in sequence.
+
+        One special hook is the rh_learn_proprio function which consists of a
+        list of hooked functions implementing the particular learning setup. The
+        rh_learn_proprio function is looped over numstep times.
+        """
+        # FIXME: obsolete
         if mode == "type01_state_prediction_error":
             self.run_hooks["hook01"] = self.rh_state_prediction_error
[…]

### 2017-06-06 active_inference_basic.py
fixed all modes for basic operation

@@ -50,7 +50,6 @@ SAVEPLOTS = True
 # DONE: modify this code to use the GMM model from actinf_models.py (ca. 20161210)
 
 modes = [
-    "m2_prediction_errors",
     "test_models",
     "type01_state_prediction_error",
     "type02_state",
@@ -58,6 +57,9 @@ modes = [
     "type03_goal_prediction_error",
     "type04_ext_prop",
     "type05_multiple_models",
+    "m2_prediction_errors",
+    "basic_operation_1D_M1",
+    "plot_system",
 ]
 
 actinf_environments = [
@@ -1085,8 +1087,8 @@ class ActiveInferenceExperiment(object):
         # print X_accum
         X = X_accum
         # X's and pred's indices now mean: slowest: goal, e1, e2, fastest: e3
+        self.X_model_sweep = X.copy()
         return X
-        # self.X_model_sweep = X.copy()
         # print "self.X_model_sweep.shape", self.X_model_sweep.shape
 
     def rh_model_sweep(self):
@@ -1098,7 +1100,7 @@ class ActiveInferenceExperiment(object):
             pred.append(self.mdl.predict(self.X_model_sweep[i]))
         pred = np.array(pred)
         print "pred.shape", pred.shape
-        # self.Y_model_sweep = pred
+        self.Y_model_sweep = pred
         return pred
                 
     def rh_model_plot(self):
@@ -1131,15 +1133,15 @@ class ActiveInferenceExperiment(object):
         pl.ioff()
 
         # generate model input sweep as meshgrid
-        self.X_model_sweep = self.rh_model_sweep_generate_input_grid()
+        X_accum = self.rh_model_sweep_generate_input_grid()
         print "map_model_type03_goal_prediction_error self.X_model_sweep = %s" % (self.X_model_sweep.shape,)
 
         # execute the model on the sweep meshgrid
-        self.Y_model_sweep = self.rh_model_sweep()
+        pred = self.rh_model_sweep()
         print "map_model_type03_goal_prediction_error self.Y_model_sweep = %s" % (self.Y_model_sweep.shape,)
 
-        X_accum = self.X_model_sweep
-        pred = self.Y_model_sweep
+        # X_accum = self.X_model_sweep
+        # pred = self.Y_model_sweep
         numgrid = self.Y_model_sweep.shape[0]
         
         # ############################################################

### 2017-06-06 active_inference_basic.py
more fixes

@@ -1095,7 +1095,7 @@ class ActiveInferenceExperiment(object):
         assert hasattr(self, "X_model_sweep")
         pred = []
         for i in range(self.X_model_sweep.shape[0]): # loop over start states
-            print "trying a predict with X = %d, %s" % (i, self.X_model_sweep.shape)
+            print "trying a predict with X[%d] = %s" % (i, self.X_model_sweep.shape)
             # this might go wrong with different models
             pred.append(self.mdl.predict(self.X_model_sweep[i]))
         pred = np.array(pred)

### 2017-06-08 active_inference_basic.py
getting actinf into clean state before merge

@@ -8,6 +8,7 @@ import pylab as pl
 import matplotlib.gridspec as gridspec
 import pandas as pd
 
+import explauto
 from explauto import Environment
 from explauto.environment import environments
 from explauto.environment.pointmass import PointmassEnvironment
@@ -301,14 +302,19 @@ class ActiveInferenceExperiment(object):
             self.run = self.run_type05_multiple_models
 
         elif mode == "m2_prediction_errors":
-            """active inference / predictive coding: first working, most basic version,
-            proprioceptive only
+            """active inference / predictive coding
 
-            goal -> goal state prediction -> goal/state error -> update forward model"""
+            most basic version, proprioceptive only
 
+            goal -> goal state prediction -> goal/state error -> update forward model
+            """
+
+            # only works for pointmass?
+            assert isinstance(self.environment, PointmassEnvironment), "Need PointmassEnvironment, not %s" % (self.environment,)
+            
             # pimp environment
-            self.environment.motor_aberration["type"] = "sin"
-            self.environment.motor_aberration["coef"] = np.random.uniform(-1.0, 1.0, self.odim) # -0.7
+            self.environment.motor_aberration["type"] = "linsin"
+            self.environment.motor_aberration["coef"] = 3.0 # np.random.uniform(-1.0, 1.0, self.odim) # -0.7
             
             # learning loop hooks
             self.rh_learn_proprio_hooks["hook04"] = self.lh_sample_discrete_uniform_goal
@@ -372,7 +378,7 @@ class ActiveInferenceExperiment(object):
         
         funcdict = OrderedDict()
         # create sweep input data
-        funcdict["hook01"] = self.rh_model_sweep_generate_input_grid
+        funcdict["hook01"] = self.rh_model_sweep_generate_input_grid_a
         # sweep and create output data
         funcdict["hook02"] = self.rh_model_sweep
         # plot result
@@ -456,9 +462,13 @@ class ActiveInferenceExperiment(object):
 
     # system sweep hooks
     def rh_system_generate_sweep_input(self):
-        """generate system inputs for grid sweep"""
+        """ActiveInferenceExperiment.rh_system_generate_sweep_input
+
+        generate system inputs on a grid to sweep the system
+        """
+        
[…]

### 2017-06-08 actinf_models.py
model testing for all models

@@ -163,6 +163,11 @@ class ActInfOTLModel(ActInfModel):
         return np.array(pred).reshape((1, self.odim))
         
     def fit(self, X, y):
+        """ActInfOTLModel.fit
+
+        Fit model to data X, y
+        """
+        
         if X.shape[0] > 1: # batch of data
             return self.fit_batch(X, y)
         
@@ -185,7 +190,7 @@ class ActInfOTLModel(ActInfModel):
 
     def fit_batch(self, X, y):
         for i in range(X.shape[0]):
-            self.fit(X[i], y[i])
+            self.fit(X[[i]], y[[i]])
         
     def save(self, filename):
         otlmodel_ = self.otlmodel
@@ -257,7 +262,7 @@ class ActInfSOESGP(ActInfOTLModel):
                     False, self.kernel_params, self.noise, self.epsilon,
                     self.capacity, self.random_seed)
         im = res_input_matrix_random_sparse(self.idim, self.res_size, 0.2)
-        print("im", type(im))
+        # print("im", type(im))
         self.otlmodel.setInputWeights(im.tolist())
 
 ################################################################################
@@ -1316,7 +1321,12 @@ def generate_inverted_sinewave_dataset(N = 1000):
     return X,Y
 
 def test_model(args):
-    import pylab as pl
+    """actinf_models.test_model
+
+    Test the model type given in args.modelclass on data
+    """
+    
+    # import pylab as pl
 
     np.random.seed(args.seed)
     
@@ -1373,10 +1383,10 @@ def test_model(args):
 
     print("Testing model class %s, %s" % (mdlcls, mdl))
 
-    print("X.shape = %s, Y.shape = %s" % (X.shape, Y.shape))
-    
+    print("Fitting model with X.shape = %s, Y.shape = %s" % (X.shape, Y.shape))
     mdl.fit(X, Y)
     
+    print("Plotting model %s, %s" % (mdlcls, mdl))
     if args.modelclass == "HebbSOM":
         
         e_nodes, p_nodes = hebbsom_get_map_nodes(mdl, idim, odim)
@@ -1406,7 +1416,7 @@ def test_model(args):
         # pl.plot(nodes, filter_e.map.neurons[:,:,1], "ko", alpha=0.5, ms=10)
         # pl.show()
     
-    else:
+    elif args.modelclass == "GMM":
         nodes = np.array(mdl.cen_lst)
[…]

### 2017-06-08 active_inference_basic.py

@@ -567,6 +567,7 @@ class ActiveInferenceExperiment(object):
 
         # sample error gradient
         numsamples = 20
+        # was @ 50
         if i % 1 == 0:
             from sklearn import linear_model
             import sklearn
@@ -578,6 +579,7 @@ class ActiveInferenceExperiment(object):
             M_ = []
             for i in range(numsamples):
                 # S_.append(np.random.normal(self.S_prop_pred, 0.01 * self.environment.conf.m_maxs, self.S_prop_pred.shape))
+                # larger sampling range
                 S_.append(np.random.normal(self.S_prop_pred, 0.3 * self.environment.conf.m_maxs, self.S_prop_pred.shape))
                 # print "S_[-1]", S_[-1]
                 M_.append(self.environment.compute_motor_command(S_[-1]))
@@ -1780,18 +1782,18 @@ def test_models(args):
     odim = 2
     numdatapoints = 10
     
-    for aimclass in [ActInfModel, ActInfKNN, ActInfSOESGP]:
-        print("aimclass", aimclass)
+    for aimclass in [ActInfModel, ActInfKNN, ActInfSOESGP, ActInfSTORKGP, ActInfGMM, ActInfHebbianSOM]:
+        print("Testing aimclass = %s" % (aimclass,))
         aim = aimclass(idim = idim, odim = odim)
 
         X = np.random.uniform(-0.1, 0.1, (numdatapoints, 1, idim))
         y = np.random.uniform(-0.1, 0.1, (numdatapoints, 1, odim))
 
         for i in range(numdatapoints-1):
+            print("Fitting model with X = %s, Y = %s" % (X[i].shape, y[i].shape))
             aim.fit(X[i], y[i])
         y_ = aim.predict(X[i+1])
-        print("prediction error = %s" % (y_ - y[i+1]))
-
+        print("Prediction error = %s" % (y_ - y[i+1]))
     
 def main(args):
 

### 2017-06-08 actinf_models.py
more cleanup and fixing

@@ -1,12 +1,15 @@
 """Active inference project code
 
+Oswald Berthold, 2016-2017
+
 This file contains the learners which can be used as adaptive models of
 sensorimotor contexts. For forward models there are
  - nearest neighbour
  - sparse online gaussian process models powered by Harold Soh's OTL library
  - gaussian mixture model
+ - hebbian connected SOM 
 
-TODO: think about a common calling convention for all model types
+TODO: common calling convention for all model types
    - including 'predict_naive' and 'predict_full' methods that would capture
      returning confidences about the current prediction
    - other variables that might be used by the context to modulate
@@ -17,11 +20,10 @@ TODO: consistency problem when sampling from probabilistic models (gmm, hebbsom,
 issues:
  - som track residual error from map training
  - som use residual for adjusting rbf width
- - som extend sampling to sample actual prediction from gaussian with unit's mu and sigma
+ - som extend sampling to sample actual prediction from gaussian with
+   unit's mu and sigma
  - plot current / final som configuration
  - plot densities
- - 
-
 """
 
 
@@ -59,7 +61,7 @@ try:
 except ImportError, e:
     print("Couldn't import lmjohns3's kohonon SOM lib", e)
 
-model_classes = ["KNN", "SOESGP", "STORKGP", "GMM", "HebbSOM"]
+model_classes = ["KNN", "SOESGP", "STORKGP", "GMM", "HebbSOM", "all"]
         
 class ActInfModel(object):
     """Base class for active inference function approximators / regressors"""
@@ -1452,7 +1454,7 @@ if __name__ == "__main__":
 
     parser = argparse.ArgumentParser()
     parser.add_argument("-d", "--datafile",     type=str, help="datafile containing t x (dim_extero + dim_proprio) matrix ", default="data/simplearm_n1000/EP_1000.npy")
-    parser.add_argument("-m", "--modelclass",   type=str, help="Which model class to test [KNN], " + ", ".join(model_classes), default="KNN")
+    parser.add_argument("-m", "--modelclass",   type=str, help="Which model class [all] to test from " + ", ".join(model_classes), default="all")
     parser.add_argument("-n", "--numsteps",     type=int, help="Number of datapoints [1000]", default=1000)
     parser.add_argument("-ne", "--numepisodes", type=int, help="Number of episodes [10]", default=10)
     parser.add_argument("-s",  "--seed",        type=int, help="seed for RNG [0]",        default=0)

### 2017-06-08 active_inference_basic.py

@@ -1,5 +1,5 @@
 
-import argparse, cPickle, os
+import argparse, cPickle, os, sys
 from collections import OrderedDict
 from functools   import partial # mhm :)
 
@@ -51,16 +51,13 @@ SAVEPLOTS = True
 # DONE: modify this code to use the GMM model from actinf_models.py (ca. 20161210)
 
 modes = [
-    "test_models",
-    "type01_state_prediction_error",
-    "type02_state",
-    "type03_1_prediction_error",
-    "type03_goal_prediction_error",
-    "type04_ext_prop",
-    "type05_multiple_models",
-    "m2_prediction_errors",
-    "basic_operation_1D_M1",
+    "m1_goal_error_1d",
+    "m1_goal_error_nd",
+    "m1_goal_error_nd_e2p",
+    "m2_error_nd",
+    "m2_error_nd_ext",
     "plot_system",
+    "test_models", # simple model test
 ]
 
 actinf_environments = [
@@ -74,7 +71,7 @@ def gaussian(m, s, x):
     return 1/(s*np.sqrt(2*np.pi)) * np.exp(-0.5*np.square((m-x)/s))
 
 class ActiveInferenceExperiment(object):
-    def __init__(self, mode = "type01_state_prediction_error",
+    def __init__(self, mode = "m1_goal_error_nd",
                  model = "knn", numsteps = 1000, idim = None,
                  environment_str = "simplearm",
                  goal_sample_interval = 50, e2pmodel = None,
@@ -93,6 +90,8 @@ class ActiveInferenceExperiment(object):
                 
         self.saveplots = saveplots
 
+        self.environment_str = environment_str
+        
         # intialize robot environment
         if environment_str == "simplearm":
             self.environment = Environment.from_configuration('simple_arm', 'low_dimensional')
@@ -225,15 +224,8 @@ class ActiveInferenceExperiment(object):
         list of hooked functions implementing the particular learning setup. The
         rh_learn_proprio function is looped over numstep times.
         """
-        # FIXME: obsolete
-        if mode == "type01_state_prediction_error":
-            self.run_hooks["hook01"] = self.rh_state_prediction_error
-            
-        # FIXME: obsolete
-        elif mode == "type02_state":
-            self.run = self.rh_state
-            
-        elif mode == "type03_goal_prediction_error":
+        
+        if mode == "m1_goal_error_nd":
             """active inference / predictive coding
 
[…]
